---
title: "chi_square_independence"
author: "Karla Mona"
date: "December 8, 2025"
format:
  pdf:
    number-sections: true
    toc: true
    pdf-engine: xelatex
editor: visual
  markdown: 
    wrap: 72
---

```{r}

installed.packages("learnr", dependencies = TRUE)
installed.packages("tidyverse", dependencies = TRUE)
installed.packages("lsr", dependencies = TRUE)
installed.packages("knitr", dependencies = TRUE)
installed.packages("gapminder", dependencies = TRUE)
installed.packages("haven", dependencies = TRUE)
```

\## Theory: What is a Chi-Square Test?

Chi-Square tests assess whether observed categorical data differ significantly from what we would expect under a specific hypothesis.

### Types of Chi-Square Tests:

1.  **Goodness-of-Fit Test** – Compares one categorical variable to a theoretical distribution.
2.  **Test of Independence** – Assesses whether two categorical variables are related.

**Formula:** $$
\chi^2 = \sum \frac{(O_i - E_i)^2}{E_i}
$$ Where: - (O_i) = Observed frequency - (E_i) = Expected frequency

\`\`\`{r} \# Load Titanic dataset data("Titanic") df \<- as.data.frame(Titanic)

# Create contingency table: Class vs Survived

```{r}
tab <- xtabs(Freq ~ Class + Survived, data = df) tab
```

# Run Chi-Square test

chi_result \<- chisq.test(tab) chi_result

# Check expected counts

chi_result\$expected

# Calculate Cramer's V

n \<- sum(tab) min_dim \<- min(nrow(tab) - 1, ncol(tab) - 1)

# 1) Load your data df \<- read.csv("your_file.csv") \# or readr::read_csv("your_file.csv")

```{r}
file_path <- "../data/raw/13578-0001-Data.por"
df <-read_por(file_path)
```

# 2) Quick check: see column names and first rows

names(df) head(df)

# 3) Make sure OQ5 and OQ6 are categorical (factors)

df$OQ5 <- factor(df$OQ5) df$OQ6 <- factor(df$OQ6)

# Optional: remove rows with missing values in these two variables

df \<- subset(df, !is.na(OQ5) & !is.na(OQ6))

# 4) Build contingency table and run Chi-square

tab \<- table(df$OQ5, df$OQ6) tab addmargins(tab)

chi_result \<- chisq.test(tab) chi_result

# 5) Check expected counts (assumption)

chi_result\$expected

# 6) Cramer's V

n \<- sum(tab) min_dim \<- min(nrow(tab) - 1, ncol(tab) - 1) cramers_v \<- sqrt(as.numeric(chi_result\$statistic) / (n \* min_dim)) cramers_v

# 7) Visualize

mosaicplot(tab, color = TRUE, main = "OQ5 × OQ6")
