---
title: "chi_square_independence"
author: "Karla Mona"
date: "December 8, 2025"
format:
  pdf:
    number-sections: true
    toc: true
    pdf-engine: xelatex
editor: visual
  markdown: 
    wrap: 72
---

```{r}

installed.packages("learnr", dependencies = TRUE)
installed.packages("tidyverse", dependencies = TRUE)
installed.packages("lsr", dependencies = TRUE)
installed.packages("knitr", dependencies = TRUE)
installed.packages("gapminder", dependencies = TRUE)
installed.packages("haven", dependencies = TRUE)
```

\## Theory: What is a Chi-Square Test?

Chi-Square tests assess whether observed categorical data differ significantly from what we would expect under a specific hypothesis.

### Types of Chi-Square Tests:

1.  **Goodness-of-Fit Test** – Compares one categorical variable to a theoretical distribution.
2.  **Test of Independence** – Assesses whether two categorical variables are related.

**Formula:** $$
\chi^2 = \sum \frac{(O_i - E_i)^2}{E_i}
$$ Where: - (O_i) = Observed frequency - (E_i) = Expected frequency

# Load Titanic dataset data("Titanic") df \<- as.data.frame(Titanic)

# Create contingency table: Class vs Survived


```{r}
# Load Titanic dataset
data("Titanic")
df <- as.data.frame(Titanic)

# Create contingency table: Class vs Survived
tab <- xtabs(Freq ~ Class + Survived, data = df)
tab

# Run Chi-Square test

chi_result \<- chisq.test(tab) chi_result

# Check expected counts

chi_result\$expected

# Calculate Cramer's V

n \<- sum(tab) min_dim \<- min(nrow(tab) - 1, ncol(tab) - 1)

# 1) Load your data
file_path <- "../data/raw/13578-0001-Data.por"
df <- read_por(file_path)

# 2) Quick check
names(df)
head(df)

# 3) Make sure Q05 and Q06 are categorical (factors)
df$Q05 <- factor(df$Q05)
df$Q06 <- factor(df$Q06)

# 4) Remove rows with missing values in Q05 or Q06
df <- subset(df, !is.na(Q05) & !is.na(Q06))

# 5) Build contingency table
tab <- table(df$Q05, df$Q06)
addmargins(tab)

# 6) Chi-square test
chi_result <- chisq.test(tab)
chi_result

# 7) Check expected counts
chi_result$expected

# 8) Cramer's V
n <- sum(tab)
min_dim <- min(nrow(tab) - 1, ncol(tab) - 1)
cramers_v <- sqrt(as.numeric(chi_result$statistic) / (n * min_dim))
cramers_v

# 9) Visualization
mosaicplot(tab, color = TRUE, main = "Q05 x Q06")