---
title: "WorkingWithData"
format: html
---

```{r}

#Load libraries

library(tidyverse)
library(haven)
library(readxl)
library(labelled)
codebook <- look_for(df)
install.packages("descr")
library(descr)



```

```{r load-data}

#Load dataset 

file_path <- "../data/raw/13578-0001-Data.por"
df <-read_por(file_path)

head(df)
```

```{r}
colnames(df)
```

#Explore Data

```{r summary}

summary(df)
```

```{r glimpse}

glimpse(df)
```

#Analize Data

```{r freq-context}
freq(as.ordered(df$OQ6), plot = FALSE)
```

#0Q6 as alcohol related context (empty beer bottles on sight) CONTEXT on PSTD as time informative interval, 1 is yes 2 is no.

```{r freq-context2}
freq(as.ordered(df$D53), plot = FALSE)

```

##drunk or intoxicated people on sight. 1 is yes, 2 is no.

```{r freq-context3}
freq(as.ordered(df$D15_12))
```

```{r freq-context4}
freq(as.ordered(df$OQ5), plot = FALSE)
```

{r freq-context2} freq(as.ordered(df\$D53), plot = FALSE)

#non related alcohol context as D15-12, 1 is no drinking establishment, and 2 means other categories

`{df$OQ6[df$OQ6 == 2] <- "no"} freq(as.ordered(df$OQ6)`

`#{df$D53[df$D53 == -99] <- "skipped"} freq(as.ordered(df$D53)`

`{df$D15[df$D15 == 1] <- "yes"} freq(as.ordered(df$D15)`

#not alcohol related context

#what the majority picked when asked of context related (the three variables 1, 2, -99)

```{r}

#this turns -99 to NA (as compared to above one without) 
#df$D53[df$D53 == -99 ] <- NA

#{df$D53[df$D53 == -99] <- NA}

freq(as.ordered(df$D53), plot = FALSE)

```

```{r}

#Data Wrangling

#colapsing OQ5 simplifying the categories

df$OQ5[df$OQ5 == 1] <- 3 #streets in construction

df$OQ5[df$OQ5 == 2 ] <- 2 # streets in poor condition 

df$OQ5[df$OQ5 == 3 | df$OQ5 == 4 | df$OQ5 == 5 ] <- 1 #streets in good condition

freq(as.ordered(df$OQ5), plot = FALSE)
```

{df\$D53\[df\$D53 == -99\] \<- "NA"} (df\$D53)

```{r}
#Data Wrangling 

#D53 drunks visible on street labeling responses

levels(df$D53) #checks the levels

df$D53 <- as.factor(df$D53) #turns to factor to enable renaming of levels

#levels(df$D53) <- ("skipped", "yes", "no") #renames the numbers to labels

#below is failed attempts to label but kept

#levels(df$D53)

#freq(as.ordered(df$D53), plot = FALSE)

#{df$D53[df$D53 == ] <- "NA"} (df$D53)
#{df$D53[df$D53 == 1] <- "yes"} (df$D53)
#{df$D53[df$D53 == 2] <- "no"} (df$D53)
#{df$D53[df$D53 == -99] <- "NA"} (df$D53)


                                          
freq(as.ordered(df$D53), plot = FALSE)

```

```{r}
#Assignment 5 univariate 

# elaborate

ggplot(data = df, aes(x = D53)) +  
  geom_bar(fill = "skyblue", color = "black") + 
  ggtitle("drunken intoxicated people on block face") +
  xlab("Camera count") +
  ylab("Response Count") +
  theme_minimal()



#This graph demostrates the amount of streets where people were seen as intoxicated which information was taken from street cameras.
```

```{r}

#Assignment 5 univariate 

#elaborate 2

ggplot(data = df, aes(x = OQ5)) + 
  geom_bar(fill = "skyblue", color = "black") +
  ggtitle("Condition of the streets in Chicago") +
  xlab("Street condition") +
  ylab("Amount of streets") +
  theme_minimal()
#This graph representated the condition of the streets by categories (1=under construction), (2=very poor), (3=fair), (4=moderately good), (5=good condition)
```

```{r}

#Example/Attempt

#categorical explanatory, categorical response ====

ggplot(data=df) +

stat_summary(aes(x=as.factor(OQ5),

 y=as.factor(D53)),

 fun.y=mean, 

 geom="bar")+

 ggtitle("Descriptive Title Here")
```

```         
```

```{r}

#Bivariate graph Assignment 6

ggplot(df, aes(x = OQ5, fill = D53)) +
  geom_bar(position = "stack") +
  labs(title = "Overlap of street conditio and intoxicated people",
       x = "Street condition",
       y = "Places where intoxicated people were seen",
       fill = "Variable 2 Category") +
  theme_minimal()

#Relation between how street conditions influence potentiall surroundings with intoxicated people
```
