---
title: "WorkingWithData"
format: html
---

```{r}
library(tidyverse)
library(haven)
library(readxl)
library(labelled)
codebook <- look_for(df)


```

```{r load-data}
file_path <- file.path("..","data","data.por")

df <-read_por(file_path)

head(df)
```

```{r}
colnames(df)
```

```{r summary}

summary(df)
```

```{r glimpse}

glimpse(df)
```

```{r freq-context}
freq(as.ordered(df$OQ6), plot = FALSE)
```

#0Q6 as alcohol related context (empty beer bottles on sight) CONTEXT on PSTD as time informative interval, 1 is yes 2 is no.

```{r freq-context2}
freq(as.ordered(df$D53), plot = FALSE)

```

##drunk or intoxicated people on sight. 1 is yes, 2 is no.

```{r freq-context3}
freq(as.ordered(df$D15_12))
```

```{r freq-context4}
freq(as.ordered(df$OQ5), plot = FALSE)
```

{r freq-context2} freq(as.ordered(df\$D53), plot = FALSE)

#non related alcohol context as D15-12, 1 is no drinking establishment, and 2 means other categories

`{df$OQ6[df$OQ6 == 2] <- "no"} freq(as.ordered(df$OQ6)`

`#{df$D53[df$D53 == -99] <- "skipped"} freq(as.ordered(df$D53)`

`{df$D15[df$D15 == 1] <- "yes"} freq(as.ordered(df$D15)`

#not alcohol related context

#what the majority picked when asked of context related (the three variables 1, 2, -99)

```{r}

#this turns -99 to NA (as compared to above one without) 
#df$D53[df$D53 == -99 ] <- NA

#{df$D53[df$D53 == -99] <- NA}

freq(as.ordered(df$D53), plot = FALSE)

```

```{r}

#Data Wrangling colapsing OQ5 simplifying the categories

df$OQ5[df$OQ5 == 1] <- 3 #streets in construction

df$OQ5[df$OQ5 == 2 ] <- 2 # streets in poor condition 

df$OQ5[df$OQ5 == 3 | df$OQ5 == 4 | df$OQ5 == 5 ] <- 1 #streets in good condition

freq(as.ordered(df$OQ5), plot = FALSE)
```

```{r}
#Data Wrangling D53 drunks visible on street labeling responses

levels(df$D53) #checks the levels

df$D53 <- as.factor(df$D53) #turns to factor to enable renaming of levels

levels(df$D53) <- c("skipped", "yes", "no") #renames the numbers to labels

#below is failed attempts to label but kept

#levels(df$D53)

#freq(as.ordered(df$D53), plot = FALSE)

#{df$D53[df$D53 == ] <- "NA"} (df$D53)
#{df$D53[df$D53 == 1] <- "yes"} (df$D53)
#{df$D53[df$D53 == 2] <- "no"} (df$D53)
                                          
freq(as.ordered(df$D53), plot = FALSE)

```
